Further Explorations of \suldiox~Aqueous Adsorption:

### 2011-09-08 Cycle-Results.tex
More of Fred's final edits

Having examined the bonding coordinations and bonding behavior of \suldiox~with surface waters, we now turn to a secondary behavior of the hydrate structures that form around the surface-bound \suldiox~molecule. The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form during MD, as posited earlier in the text and depicted in Figure \ref{fig:cyclic-structures}. Only the most commonly occurring subset of the cyclic structures were analyzed based on two selection criteria for determining formation of a cyclic structure: (1) The distances between atoms must match the same bonding/distance criteria as used for determining bonding coordinations. (2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen. As noted earlier in the discussion of the graph-finding routine and the BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on the search's order of cycle discovery. The \suldiox~will be involved in other larger and more extended cyclic bonding structures beyond the first one discovered via the BFS. The larger and more extended cyclic structures involving more waters affect the behavior of the hydrogen-bonding network of the water surface. However, we focus here only on the smallest cycles involving the \suldiox~as they most affect the \suldiox~bonding and hydration.

@@

The plot in Figure \ref{fig:cyclic-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures. Each data point shows a percentage of the MD trajectories in which the \suldiox~was a member of a cyclic structure, for different numbers of cyclic waters (up to 4). The two tallest data points, left-most in the plot, show the overall time spent in all types of cyclic structures. Clearly the hot system \suldiox~spends more time in a cyclic structure than at the cold temperature. It is remarkable that the difference of the times spent in a cyclic structure between the two temperatures is 15\% (42\% cold, 57\% hot). The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such. Thus, in addition to having earlier found the most likely bonding coordination during the simulated life of \suldiox, we have also found that the hydrates of the \suldiox~form cyclic structures for much of the time while bound to the water surface.

Now we look at the different types of cyclic hydrates, distinguished by the number of waters involved in the bonding structure. In Figure \ref{fig:cyclic-breakdown} the single and double water cycles are the least frequently encountered structures, accounting for less than 10\% of both cold and hot temperature simulation times. Formation of the single-water type is likely energetically unfavorable because of the proximity of the water to the \suldiox~required to form the bonding cycle. The double-water structure was one of two types of clusters proposed in a previous computational work as a candidate structure contributing to the overall IR spectrum of surface-bound \suldiox.\cite{Baer2010} In the static and geometry-optimized cluster calculations, lacking the extended water structure or bonding from waters external to the hydrate, both the double and triple types appear equally likely to form. However, the MD simulations here have introduced many waters into a dynamic environment allowing for extended bonding networks, and the results show clearly that the double-water cyclic hydrate is formed much less often (less than 5\% at both temperatures) than the triple-water form.

The results for triple and quadruple-water structures show that larger hydrate cycles are favored at higher temperatures. Although the higher number of hydrating waters ($>$ 4) are not shown, those contribute minimally to the overall distribution. The majority of the cyclic hydrates are formed with three waters in the triple-water type. This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules (described earlier and shown in Figure \ref{fig:type-3-varieties}). The triple-water type cycle results were broken down into contributions from the type-A and type-B triple-water structures. The two triple cycle structures are depicted on the right side of Figure \ref{fig:cyclic-breakdown}, along with the plots of their contributions to the overall distribution. It is notable that each temperature has a different dominant type of triple-water cyclic structure. The cold system forms more of the type-B, and the hotter system forms primarily type-A.

@@

We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions. The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states. $f(t)$ may then be used to calculate statistics about the cycle lifespans. A representative portion of a cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time. The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each. The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration. The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/cycles/cycle-debouncing-small.png}
		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures. A debouncing procedure was developed (described above in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure. A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure. A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
		\label{fig:debouncing}
	\end{center}
\end{figure}

The distribution of cycle lifetimes (i.e. contiguous spans of time spent with $f(t) = 1$) was determined. The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations. The most frequent lifespan for both temperature data sets lasts between 0-1 ps. This accounts for the vast majority of cyclic hydrates found (approximately 95\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time. Even with the debouncing procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps. The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend also extending to the longer lifespans as well. The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 1 ps. All of the distribution shows only a $<1.5$\% difference between cold and hot cyclic lifetimes. Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time. Up to 3 ps, the cold temperature cycles show a very slight population increase above the hot temperature cycles. Above 4 ps, most of the cycles that form are found in the hot system. The 8 ps cycles are notable in that they form  for just under half of an entire single simulated trajectory.

>>>>

Having examined the bonding coordinations and bonding behavior of \suldiox~with surface waters, we now turn to a secondary behavior of the hydrate structures that form around the surface-bound \suldiox~molecule. The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form, as posited earlier in the text and depicted in Figure \ref{fig:cyclic-structures}. Only the most commonly occurring subset of the cyclic structures were analyzed based on two selection criteria: (1) The distances between atoms must match the same bonding/distance criteria as used for determining bonding coordinations. (2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen. As noted earlier in the discussion of the graph-finding routine and the BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on the search's order of cycle discovery. The \suldiox~will be involved in other larger and more extended cyclic bonding structures beyond the first one discovered via the BFS. The larger and more extended cyclic structures involving more waters affect the behavior of the hydrogen-bonding network of the water surface. We focus here only on the smallest cycles involving the \suldiox~as they most affect the \suldiox~bonding and hydration.

@@

The plot in Figure \ref{fig:cyclic-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures. Each data point shows a percentage of the MD trajectories in which the \suldiox~was a member of a cyclic structure, for different numbers of cyclic waters (up to 4). The two tallest data points, left-most in the plot, show the overall time spent in all types of cyclic structures. Clearly the hot \suldiox~spends more time in a cyclic structure than at the cold temperature. It is remarkable that the time spent in a cyclic structure shows a 15\% difference (42\% cold, 57\% hot). The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such. Thus, in addition to having earlier found the most likely bonding coordination during the simulated life of \suldiox, we have also found that the hydrates of the \suldiox~form cyclic structures for much of the time while bound to the water surface.

Now we look at the different types of cyclic hydrates, distinguished by the number of waters involved in the bonding structure. In Figure \ref{fig:cyclic-breakdown} the single and double water cycles are the least frequently encountered structures, accounting for less than 10\% of both cold and hot temperature simulation times. Formation of the single-water type is likely energetically unfavorable because of the proximity of the water to the \suldiox~required to form the bonding cycle. The double-water structure was one of two types of clusters proposed in a previous computational work as a candidate structure contributing to the overall IR spectrum of surface-bound \suldiox.\cite{Baer2010} In those static and geometry-optimized cluster calculations, lacking the extended water structure or bonding from waters external to the hydrate, both the double and triple types appear equally likely to form. However, the MD simulations here have introduced many waters into a dynamic environment allowing for extended bonding networks, and the results show clearly that the double-water cyclic hydrate is formed much less often (less than 5\% at both temperatures) than the triple-water form.

The results for triple and quadruple-water structures show that larger hydrate cycles are favored at higher temperatures. Although the higher number of hydrating waters ($>$ 4) are not shown, those contribute minimally to the overall distribution. The majority of the cyclic hydrates are formed with three waters in the triple-water type. This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that, of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules (described earlier and shown in Figure \ref{fig:type-3-varieties}). The triple-water type cycle results were broken-down into contributions from the type-A and type-B triple-water structures. The two triple cycle structures are depicted on the right side of Figure \ref{fig:cyclic-breakdown}, along with the plots of their contributions to the overall distribution. It is notable that each temperature has a different dominant type of triple-water cyclic structure. The cold system forms more of the type-B, and the hotter system forms primarily type-A.

@@

We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions. The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states. $f(t)$ may then be used to calculate statistics about the cycle lifespans. A representative portion of a cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time. The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each. $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration. The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/cycles/cycle-debouncing-small.png}
		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures. A debouncing procedure was developed (described in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure. A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure. A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
		\label{fig:debouncing}
	\end{center}
\end{figure}

The distribution of cycle lifetimes (i.e. contiguous spans of time spent with $f(t) = 1$) was determined. The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations. The most frequent lifespan for both temperature data sets lasts between 0-1 ps. This accounts for the vast majority of cyclic hydrates found (approximately 95\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time. Even with the debouncing procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps. The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend extending to the longer lifespans as well. The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 1 ps. All of the distribution shows only a $<1.5$\% difference between cold and hot cyclic lifetimes. Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time. Up to 3 ps, the cold temperature cycles show a very slight population increase above the hot temperature cycles. Above 4 ps, most of the cycles that form are found in the hot system. The 8 ps cycles are notable in that they form for just under half the length of one of the simulated trajectories.

### 2011-09-08 Introduction.tex
More of Fred's final edits

One of Chemistry's many remaining ``black box'' events is the adsorption of gas molecules onto a water surface. Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions. How does gas initially bind to a water surface, and what steps are involved in the subsequent adsorption? How does an unbound gas molecule near a water surface affect the water to which it will bind? What is the structure of hydrating waters in the surface region, and how does a hydrated solute molecule behave differently than as a gas? Experiments to address these questions provide valuable information, but can never fully characterize these microscopic events and behaviors. However, these systems can be fully characterized computationally, and when coupled to the previous experimental work can provide a much more complete picture of gaseous adsorption to aqueous surfaces.

As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems. \cite{Boniface2000,Jayne1990,Johns2011,Heber1997,Faloona2009,Clegg2001} Its molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions. A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in interfacial water regions.

In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on water. We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the variability of the \suldiox~hydrate structures, and the behaviors of the water surface. The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated. In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas. Understanding the orientational behavior of molecules in the aqueous interfacial region during adsorption was a necessary first step to understanding the specific details of gas binding and surface behavior.

@@

We believe this to be the first temperature study using quantum MD to study the binding of small gas molecules on a water surface. We show how temperature affects the bonding behavior of the surface adsorbed \suldiox~to neighboring waters, and propose a sequential binding mechanism for \suldiox~adsorbing to a water surface. We also examine \suldiox~binding behavior when bound to the surface waters. Lastly, our analysis demonstrates an extended bonding structure of \suldiox~hydrates, as they are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.

>>>>

One of Chemistry's many remaining ``black box'' events is the adsorption of gas molecules onto a water surface. Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions. How does gas initially bind to a water surface, and what steps are involved in the subsequent adsorption? How does an unbound gas molecule near a water surface affect the water to which it will bind? What is the structure of hydrating waters in the surface region, and how does a hydrated solute molecule behave differently than as a gas? Experiments to address these questions provide valuable information, but can never fully characterize microscopic events and behaviors. However, these systems can be fully characterized computationally, and when coupled to the previous experimental work can provide a much more complete picture of gaseous adsorption to aqueous surfaces.

As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems. \cite{Boniface2000,Jayne1990,Johns2011,Heber1997,Faloona2009,Clegg2001} Its molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions. A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in water's surface region.

In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on water. We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the variability of the \suldiox~hydrate structures, and the behavior of the water surface. The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated. In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas. Understanding the orientational behavior of molecules in the aqueous interfacial region during adsorption was a necessary first step to understanding the specific details of gas-binding and surface behavior.

@@

We believe this to be the first temperature study using quantum MD to study the binding of small gas molecules on a water surface. We show how temperature affects the bonding behavior of the surface-adsorbed \suldiox~to neighboring waters, and propose a sequential binding mechanism for \suldiox~adsorbing to a water surface. We also examine \suldiox~binding behavior when bound to the surface waters. Lastly, our analysis demonstrates an extended bonding structure of \suldiox~hydrates, as they are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.
